---
title: "GRIP SPARK FOUNDATION TASK 3"
output: html_notebook
author: ASWIN.B
task:   Exploratory Data analysisp-Retail 
---

Importing Required Libraries
```{r}
library(heatmap3)
library(ggthemes)
library(sf)
library(raster)
library(dplyr)
library(spData)
library(spDataLarge)
library(tmap)
library(leaflet) 
library(plotly)
library(dplyr)
library(readr)
library(Hmisc)
library(corrplot)
library(reshape2)
library("PerformanceAnalytics")
library(ggplot2)
library(tidyverse)
library(GGally)
library(plotrix)
library(lattice)
library(maps)
library(ggthemes)
library(tmaptools)
library(htmltools)
library(mapproj)
library(ggmap)
library(mapdata)
library(leaflet.extras)
library(patchwork)
library(Rmisc)
library(gridExtra)

```

Reading the dataset 

```{r}
library(readr)
Task3 <- read_csv("Task3.csv")
View(Task3)
```

Analyzing the data set

```{r}

Dataset=Task3
colnames(Dataset)
summary(Dataset)
nrow(Dataset)
```

Deleting the duplicate data
```{r}
di=Dataset[duplicated(Dataset),]
Dataset=Dataset[!duplicated(Dataset),]
nrow(Dataset)
```
We can see that our rows has been reduced from 9994 to 9977.

```{r}
describe(Dataset)
```

Finding covariance and Correlation of numerical columns
```{r}
Newdata=Dataset[,c(10:13)]
round(Newdata,2)
head(Newdata)
cor(Newdata)
cov(Newdata)

```

Ploting the correlation

```{r}
corrplot(cor(Newdata),method = 'circle')
corrplot(cor(Newdata), type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)
```

Heat map and correlation chart for checking correlation

```{r}
chart.Correlation(Newdata, histogram = TRUE, pch = 19)
col<- colorRampPalette(c("blue", "white", "dark red"))(25)
heatmap(x =cor(Newdata) , col = col, symm = TRUE)

```
```{r}
melted_Newdata <- melt(cor(Newdata))
head(melted_Newdata)

melted_Newdata$value= round(melted_Newdata$value,2)
ggplot(data = melted_Newdata, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile()+geom_text(aes(Var2, Var1, label = value), color = "black", size = 4)
```
Analyzing Ship mode
```{r}
count1 <- as.data.frame(table(Dataset$`Ship Mode`))
ggplot(count1, aes(x = Var1, y = Freq, fill = Var1)) +  # Plot with values on top
  geom_bar(stat = "identity") +
  geom_text(aes(label = Freq), vjust = 0)
```
Correlation of sales with respect to quantity , profits , discount
```{r}
ggpairs(Dataset, columns = 10:13, ggplot2::aes(colour=`Ship Mode`)) 
```


Box plots of Profit vs Ship mode
```{r}
ggplot(Dataset,aes(reorder(`Ship Mode`,-Profit,median),Profit))+
      geom_boxplot()+
      coord_flip()+
      scale_x_discrete("`Ship Mode`")

```
Count of different categories based on shipment
```{r}
ggplot(Dataset,aes(x=`Ship Mode`,fill= Category)) +
       geom_bar(stat="count")
```
Segment Wise Analysis
```{r}
Total_Segment = aggregate(c(Dataset['Quantity'],Dataset['Sales'],Dataset['Discount'],Dataset['Profit']), by =Dataset['Segment'],FUN = sum)
View(Total_Segment) 
```
Count of Different segments
```{r}
count2 <- as.data.frame(table(Dataset$Segment))
ggplot(count2, aes(x = Var1, y = Freq, fill = Var1)) +  # Plot with values on top
  geom_bar(stat = "identity") +
  geom_text(aes(label = Freq), vjust = 0)
```
Correlation of numerical columns with respect to segments
```{r}
ggpairs(Dataset, columns = 10:13, ggplot2::aes(colour=Segment))
```
Box plots of different segments based on profits
```{r}
ggplot(Dataset,aes(reorder(Segment,-Profit,median),Profit))+
  geom_boxplot()+
  coord_flip()+
  scale_x_discrete("Segment")
```

Quantity vs Profit
```{r}
plot(Dataset$Quantity,
     Dataset$Profit,
     pch = 18,
     col = 2,
     cex = 1.2,
     xlab = "Quantity",
     ylab = "Profit")
cor(Dataset$Quantity,Dataset$Profit)

```

sales vs profit
```{r}
plot(Dataset$Sales,
     Dataset$Profit,
     pch = 17,
     col = 3,
     cex = 1.2,
     xlab = "Sales",
     ylab = "Profit")
cor(Dataset$Sales,Dataset$Profit)
```
Category Wise Analysis

```{r}
Obs_cat = aggregate(c(Dataset['Quantity'],Dataset['Sales'],Dataset['Discount'],Dataset['Profit']), by =Dataset[8],FUN = sum)
Obs_cat$PP=round(((Obs_cat$Profit/sum(Obs_cat$Profit))*100),2)
View(Obs_cat)
```

Count of categories based on subcategories
```{r}
ggplot(Dataset,aes(x=Category,fill= `Sub-Category`)) +
  geom_bar(stat="count")
```
Correlation associated with different categories
```{r}
ggpairs(Dataset, columns = 10:13, ggplot2::aes(colour=Category))
```

Total No of different categories 
```{r}
count3 <- as.data.frame(table(Dataset$Category))
ggplot(count3, aes(x = Var1, y = Freq, fill = Var1)) +  # Plot with values on top
  geom_bar(stat = "identity") +
  geom_text(aes(label = Freq), vjust = 0)
```

Box plot of profit vs category
```{r}
ggplot(Dataset,aes(reorder(Category,-Profit, median),Profit))+
  geom_boxplot(width=0.5)+
  coord_flip()+
  scale_x_discrete("Category")
```


Profit percent of Different Categories
```{r}
ggplot(Obs_cat, aes(x=Category, y=PP, fill=Category)) + 
  geom_bar(stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=PP-sd(PP), ymax=PP+sd(PP)), width=.1,
                position=position_dodge(.9))

```
Subcategory Wise Analysis

```{r}
Obs_Subcat = aggregate(c(Dataset['Quantity'],Dataset['Sales'],Dataset['Discount'],Dataset['Profit']), by =Dataset[9],FUN = sum)
Obs_Subcat$PG=round(((Obs_Subcat$Quantity/sum(Obs_Subcat$Quantity))*100),0)
Obs_Subcat$PP=round(((Obs_Subcat$Profit/sum(Obs_Subcat$Profit))*100),2)
View(Obs_Subcat)
```

Total quantity sold in different categories
```{r}
pie = ggplot(Obs_Subcat, aes(x="", y=-PG, fill=reorder(`Sub-Category`,-PG))) + geom_bar(stat="identity",color="black")

# Convert to pie (polar coordinates) and add labels
pie = pie + coord_polar("y") + 
      theme_void()

# Add color scale (hex colors)
#pie = pie + scale_fill_manual(values=c("#55DDE0", "#33658A", "#2F4858", "#F6AE2D", "#F26419", "#999999")) 

# Remove labels and add title
pie = pie + labs(x = NULL, y = NULL, fill = NULL, title = "Quantity sold in different categories")

# Tidy up the theme
pie = pie + theme_classic() + theme(axis.line = element_blank(),
                                    axis.text = element_blank(),
                                    axis.ticks = element_blank(),
                                    plot.title = element_text(hjust = 0.5, color = "#666666"))
pie
```

Analysis of Subcategory based on profit
```{r}
ggplot(Obs_Subcat, aes(x=`Sub-Category`, y=Profit, fill=`Sub-Category`)) + 
  geom_bar(stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=Profit-sd(Profit), ymax=Profit+sd(Profit)), width=.1,
                position=position_dodge(.9))+ theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

Analysis of Technology Category
```{r}
Obs_TOT = aggregate(c(Dataset['Quantity'],Dataset['Sales'],Dataset['Discount'],Dataset['Profit']), by =c(Dataset[8],Dataset[9]),FUN = sum)
Obs_Tech <- Obs_TOT %>%
  filter(Category == "Technology") %>%
  select(`Sub-Category`,Quantity, Sales, Profit)
Obs_Tech$PS=Obs_Tech$Profit/Obs_Tech$Sales
Obs_Tech$PP=round((Obs_Tech$PS/sum(Obs_Tech$PS))*100,2)
View(Obs_Tech)
```
Profit percent of Technology Category
```{r}
labels_new<-paste(Obs_Tech$`Sub-Category`,Obs_Tech$PP)
labels_new

#concatenates the above output with the '%' symbol 
final_labels<-paste(labels_new,'%',sep = "")
final_labels
pie3D(Obs_Tech$PP,labels=final_labels,explode=0.2,radius = 2.5,
      main="Profit percent of Technology subcategory ",shade = 0.5,labelcex = 0.8)
```
Analysis of furniture category
```{r}
Obs_Fur <- Obs_TOT %>%
  filter(Category == "Furniture") %>%
  select(`Sub-Category`,Quantity, Sales, Profit)
Obs_Fur$PS=Obs_Fur$Profit/Obs_Fur$Sales
Obs_Fur$PP=round((Obs_Fur$PS/sum(Obs_Fur$PS))*100,2)
View(Obs_Fur)
```


Sales percent of furniture category
```{r}
labels_new1<-paste(Obs_Fur$`Sub-Category`,round((Obs_Fur$Sales/sum(Obs_Fur$Sales))*100,2))
labels_new1

#concatenates the above output with the '%' symbol 
final_label1<-paste(labels_new1,'%',sep = "")
final_label1
pie3D(Obs_Fur$Sales,labels=final_label1,explode=0.2,radius = 2.5,
      main="Sales percent of Furniture subcategory ",shade = 0.5,labelcex = 0.8)
```

Error plot of furniture category

```{r}

ggplot(Obs_Fur, aes(x=`Sub-Category`, y=Profit, fill=`Sub-Category`)) + 
  geom_bar(stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=Profit-sd(Profit), ymax=Profit+sd(Profit)), width=.1,
                position=position_dodge(.9))

```

Sales percent of office Category
```{r}
Obs_Off <- Obs_TOT %>%
  filter(Category == "Office Supplies") %>%
  select(`Sub-Category`,Quantity, Sales, Profit)
Obs_Off$PS=Obs_Off$Profit/Obs_Off$Sales
Obs_Off$PP=round((Obs_Off$PS/sum(Obs_Off$PS))*100,2)
View(Obs_Off)

pie = ggplot(Obs_Off, aes(x="", y=-Sales, fill=reorder(`Sub-Category`,-Sales))) + geom_bar(stat="identity",color="black")

# Convert to pie (polar coordinates) and add labels
pie = pie + coord_polar("y") + 
  theme_void()

# Remove labels and add title
pie = pie + labs(x = NULL, y = NULL, fill = NULL, title = "Sales percent of Office category")

# Tidy up the theme
pie = pie + theme_classic() + theme(axis.line = element_blank(),
                                    axis.text = element_blank(),
                                    axis.ticks = element_blank(),
                                    plot.title = element_text(hjust = 0.5, color = "#666666"))
pie
```

Error plot of subcategories and their respective profits
```{r}
ggplot(Obs_Off, aes(x=`Sub-Category`, y=Profit, fill=`Sub-Category`)) + 
  geom_bar(stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=Profit-sd(Profit), ymax=Profit+sd(Profit)), width=.1,
                position=position_dodge(.9))


```
Region Wise Analysis
```{r}
Quant_Region= aggregate(c(Dataset['Sales'],Dataset['Profit']), by =Dataset[7],FUN = sum)
View(Quant_Region)
Quant_Region$PS=Quant_Region$Profit/Quant_Region$Sales
Quant_Region$PP=round((Quant_Region$PS/sum(Quant_Region$PS))*100,2)

```
Profit percent per sales taken region wise
```{r}
labels_new3<-paste(Quant_Region$Region,Quant_Region$PP)
labels_new3

#concatenates the above output with the '%' symbol 
final_label3<-paste(labels_new3,'%',sep = "")
final_label3
pie3D(Quant_Region$PP,labels=final_label3,explode=0.2,radius = 2.5,
      main="Profit percent per sales regionwise",shade = 0.5,labelcex = 0.8)

```
Density graph of sales in different region for respective categories
```{r}
ggplot(data=Dataset, aes(x=Category, fill=Region)) +
  geom_density() +
  facet_grid(Region~.)
```

RegionWise analysis of different Categories
```{r}
ggplot(data=Dataset, aes(x=Category)) +
  geom_bar() +
  facet_wrap(.~Dataset$Region, nrow=2)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
Region Wise sales of Subcategory items
```{r}
ggplot(data=Dataset, aes(x=`Sub-Category`)) +
  geom_bar() +
  facet_wrap(.~Dataset$Region, nrow=2)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


```
State Wise analysis
```{r}
Quant_State= aggregate(c(Dataset['Quantity'],Dataset['Sales'],Dataset['Discount'],Dataset['Profit']), by =Dataset[5],FUN = sum)
View(Quant_State)
```
Sales- state wise
```{r}
us_states <- map_data("state")
head(us_states)
Quant_State$region <- tolower(Quant_State$State)
us_states_sales <- left_join(us_states, Quant_State)
View(us_states_sales)


p0 <- ggplot(data = us_states_sales,
             mapping = aes(x = long, y = lat, group = group, fill = Sales))

p1 <- p0 + geom_polygon(color = "gray90", size = 0.1) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) 

p1 + labs(title = "Sales in differnt state") + theme_map() + labs(fill = "Sales")

p2 <- p1 + scale_fill_gradient(low = "white", high = "dark blue") +
  labs(title = "Sales in different state") 
us_sales_map =p2 + theme_map() + labs(fill = "Sales")
tmap_mode("view")
us_sales_map
```
City Wise Analysis
```{r}
Quant_City= aggregate(c(Dataset['Quantity'],Dataset['Sales'],Dataset['Discount'],Dataset['Profit']), by =Dataset[4],FUN = sum)
MaxSale_City=head(arrange(Quant_City,desc(Sales)), n = 10)
MinSale_City=head(arrange(Quant_City,Sales), n = 10)

MaxProf_City=head(arrange(Quant_City,desc(Profit)), n=10)
MinProf_City=head(arrange(Quant_City,Profit), n = 10)
MinProf_City
MaxProf_City

```
Finding city with maximum profit and their corresponding sales
```{r}
value1 <-c(MaxProf_City$Sales,MaxProf_City$Profit)
type1 <-c(rep("Sales", 10), rep("Profit", 10))
mydata <-data.frame(MaxProf_City$City, value1)
p <-ggplot(mydata, aes(MaxProf_City.City, value1))
p +geom_bar(stat = "identity", aes(fill = type1), position = "dodge")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
Finding city with minimum profit and their corresponding sales
```{r}
value2 <-c(MinProf_City$Sales,MinProf_City$Profit)
type2 <-c(rep("Sales", 10), rep("Profit", 10))
mydata2 <-data.frame(MinProf_City$City, value2)
p <-ggplot(mydata2, aes(MinProf_City.City, value2))
p +geom_bar(stat = "identity", aes(fill = type2), position = "dodge")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

Analysis the Sales of Some States on The Basics of High Profit, Medium and Loss
```{r}
myplots <- list() 
i=1
state_data_viewer=function(states)
  {
  product_data =Dataset %>%
  group_by(State) 
 
   for(state in states)
    {
    cp <- product_data %>%
    filter(State == state ) %>%
    select(Category,`Sub-Category`,Quantity, Sales, Profit)%>%group_by(Category)
   
     for(cat in c('Furniture', 'Office Supplies', 'Technology'))
      {
      
      kp <- cp %>% filter(Category == cat) %>% select(`Sub-Category`,Profit)
      kp2= aggregate(kp["Profit"],by= kp[2], FUN = sum)
      b<-ggplot(kp2, aes(x=`Sub-Category`, y=Profit, fill=`Sub-Category`)) +
      geom_bar(stat="identity")+theme_minimal()+ theme(axis.text.x = element_text(angle = 90, hjust = 1))+ ggtitle(state)+xlab(cat)
      myplots[[i]]=b
      i=i+1
     
     }
    }  
  multiplot(plotlist=myplots,cols = 3)
}
states = c('New York','California','Arizona')
state_data_viewer(states)
```

Observations and Conclusions: 
State wise analysis of Sales
1. California States Has Fist Highest Number of Sales
2. New York States Has Second Highest Number of Sales
3. Philadelphia suffers maximum losses and New York city earns maximum profits.
4. TEXAS offers maximum discount but the sales are less and thus ends up in losses.
5. New York, California has the maximum sales and offers a little discount making large profits.
6. States such as West Virginia offers no discount and ends up in losses.
7. Most no. of buyers are from New York and California.  
   Just focus on the states that has weak sales and provide the discount in the intial stage to uplift the sales

We can see from the correlation chart that Sales and profit are moderately correlated, discount and profit are Negatively Correlated, Quantity and profit are less correlated.

Ship mode is generally of below average 
Furnishers have the least deal, expanding the deal amount would acquire more benefits. 
Tables selling cost ought to be expanded to give benefits. 
The clasp is the least enjoyed office item and has the base deals. 
The cost of provisions ought to be increment for more noteworthy benefits 
West Region made the most benefit and Central made the less Profit.

We have 3 regions of concern, Bookcases, Supplies and Tables as we are confronting a misfortune on the deals of these things. On the deals of tables, we are confronting the most elevated misfortune. 
The least deals are in the southern locale, we can offer them more markdown. 
Deals and benefit are more in the Standard class. 
No connection among's benefit and rebate. 
We should restrict the deals of furniture and increment that of office supplies and innovation as it produces next to no benefit contrasted with its deals. 
In the sub-classes, we are confronting an enormous misfortune on tables so that ought to be limited. 
The deals and benefit in Southern and Central locale are less so we should give more motivating forces like the markdown to build deals, consequently benefit will increment.









